---
title: "Arachnid Dollo"
author: "Sabrina Pankey"
date: "2025-09-08"
output: html_document
engine: knitr
knitr:
  opts_chunk: 
    R.options:
      width: 120
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F,fig.height = 12,include=F)

library(this.path)
setwd(this.dir())

library(viridis)
#colors
present.col<-"black" #OG exists (due to inheritance or to gain)
absent.col<-"lightgray"

col.a="gray" #a=absent & not present in ancestor
col.p="#218380" #p=inherited from ancestor (previous node)
col.g="#fbb13c" #g=gain compared to previous node
col.l="#d81159" #l=lost from previous node

require(phytools)
library(phangorn)
library(dplyr)
library(tidyr)
library(ggplot2);library(ggtree)
```
```{r preptrees,include=F,eval=F}
fixnames<-function(TREE){
  TREE$tip.label[TREE$tip.label=="Hypochithonius.genome"]<-"Hypochithonius_genome"
TREE$tip.label[TREE$tip.label=="Limulus.genome"]<-"Limulus_genome"
return(TREE)
}
tr0<-read.tree("Constraint_Arachnida.treefile.tre") %>% ladderize() %>% fixnames(.)
tr1<-read.tree("Constraint_Acaromorpha.treefile.tre")  %>% ladderize() %>% fixnames(.)
tr2<-read.tree("Constraint_Cephalosomata.treefile.tre")  %>% ladderize() %>% fixnames(.)
tr3<-read.tree("Constraint_Poecilophysidea.treefile.tre")  %>% ladderize() %>% fixnames(.)
tr4<-read.tree("Constraint_Pseudos_plus_Acariformes.treefile.tre")  %>% ladderize() %>% fixnames(.)
tr5<-read.tree("Constraint_Xiph_plus_Arachnopul.treefile.tree")  %>% ladderize() %>% fixnames(.)
tr6<-read.tree("Constraint_Xiph_Ricinulei_Arachnopul.treefile.tre")  %>% ladderize() %>% fixnames(.)
tr7<-read.tree("Constraint_Acarif_Rici.treefile") %>% ladderize() %>% fixnames(.)
tr8<-read.tree("Unconstrained.treefile") %>% ladderize() %>% fixnames(.)
tr9<-read.tree("Constraint_Acarif_Opis.tre") %>% ladderize() %>% fixnames(.)
tr10<-read.tree("Constraint_Acari_take2.tree") %>% ladderize() %>% fixnames(.)

plot(tr10,cex=0.8); nodelabels(cex=0.8)
tr0<-tr0 %>% phytools::reroot(.,147,position=0.01) 
tr1<-tr1 %>% phytools::reroot(.,149,position=0.01) 
tr2<-tr2 %>% phytools::reroot(.,138,position=0.01)
tr3<-tr3 %>% phytools::reroot(.,138,position=0.01)
tr4<-tr4 %>% phytools::reroot(.,140,position=0.01)
#tr5 is rooted correctly
tr5$tip.label %in% rownames(ogdf)
tr6<-tr6 %>% phytools::reroot(.,135,position=0.01)
plot(tr6,cex=0.6); nodelabels()
tr7<-tr7 %>% phytools::reroot(.,138,position=0.01)
tr8<-tr8 %>% phytools::reroot(.,136,position=0.01)
tr8$tip.label %in% rownames(ogdf)
tr9<-tr9 %>% phytools::reroot(.,142,position=0.01)
tr10<-tr10 %>% phytools::reroot(.,146,position=0.01)

#did labels for tr3
labs<-read.csv("arachnidnodes.csv") %>% arrange(Node)
names(labs)
pdf("treeslabelled.pdf",height=11,width=8, useDingbats = F)
plot(tr0,main="Constraint_Arachnida",cex=0.6,x.lim = c(-.1, 1.5)); nodelabels(labs$Constraint_Arachnida,cex=0.6,frame="none",col="blue",adj=1)
plot(tr1,main="Constraint_Acaromorpha.treefile.tre",cex=0.6,x.lim = c(-.1, 1.5)); nodelabels(labs$Constraint_Acaromorpha,cex=0.6,frame="none",col="blue",adj=1)
plot(tr2,main="Constraint_Cephalosomata.treefile.tre",cex=0.6,x.lim = c(-.1, 1.5)); nodelabels(labs$Constraint_Cephalosomata,cex=0.6,frame="none",col="blue",adj=1)
plot(tr3,main=" Constraint_Poecilophysidea labelled",cex=0.6,x.lim = c(-.1, 1.5)); nodelabels(labs$Constraint_Poecilophysidea,cex=0.6,frame="none",col="blue",adj=1)
plot(tr4,main="Constraint_Pseudos_plus_Acariformes.treefile.tre",cex=0.6,x.lim = c(-.1, 1.5)); nodelabels(labs$Constraint_Pseudos_plus_Acariformes,cex=0.6,frame="none",col="blue",adj=1)
plot(tr5,main="Constraint_Xiph_plus_Arachnopul.treefile.tree",cex=0.6,x.lim = c(-.1, 1.5)); nodelabels(labs$Constraint_Xiph_plus_Arachnopul,cex=0.6,frame="none",col="blue",adj=1)
plot(tr6,main="Constraint_Xiph_Ricinulei_Arachnopul.treefile.tre",cex=0.6,x.lim = c(-.1, 1.5)); nodelabels(labs$Constraint_Xiph_Ricinulei_Arachnopul,cex=0.6,frame="none",col="blue",adj=1)
plot(tr7,main="Constraint_Acarif_Rici.treefile",cex=0.6,x.lim = c(-.1, 1.5)); nodelabels(labs$Constraint_Acarif_Rici,cex=0.6,frame="none",col="blue",adj=1)
plot(tr8,main="Unconstrained.treefile",cex=0.6,x.lim = c(-.1, 1.5)); nodelabels(labs$Unconstrained,cex=0.6,frame="none",col="blue",adj=1)
plot(tr9,main="Constraint_Acarif_Opis.tre",cex=0.6,x.lim = c(-.1, 1.5)); nodelabels(labs$Constraint_AcariOpiliones,cex=0.6,frame="none",col="blue",adj=1)
plot(tr10,main="Constraint_Acari_take2.tree",cex=0.6,x.lim = c(-.1, 1.5)); nodelabels(labs$Constraint_AcariParasitif,cex=0.6,frame="none",col="blue",adj=1)

dev.off()

#compare to older trees to set up nodelabels
load("../july_run/Chelicerate_og4dollo.Rdata")
phyloch::compare.phylo(tr1,tr3)
comparePhylo(tr1,tr4,plot=T)

```


```{r setupdata,include=F}
if(!file.exists("Chelicerate_og4dollo.Rdata")){
#1. prep data

labs<-read.csv("arachnidnodes.csv") %>% arrange(Node)
nodenames0<-labs$Constraint_Arachnida
nodenames1<-labs$Constraint_Acaromorpha
nodenames2<-labs$Constraint_Cephalosomata
nodenames3<-labs$Constraint_Poecilophysidea
nodenames4<-labs$Constraint_Pseudos_plus_Acariformes
nodenames5<-labs$Constraint_Xiph_plus_Arachnopul
nodenames6<-labs$Constraint_Xiph_Ricinulei_Arachnopul
nodenames7<-labs$Constraint_Acarif_Rici
nodenames8<-labs$Unconstrained
nodenames9<-labs$Constraint_AcariOpiliones
nodenames10<-labs$Constraint_AcariParasitif

treelist<-setNames(colnames(labs[-1]),nm=c("tr03","tr01","tr02","tr04","tr05","tr06","tr0","tr07","tr08","tr09","tr10"))
plot(tr1)
nodelabels(nodenames1)

plot(tr2)
nodelabels(nodenames2)


ogdf<-t(read.table("../july_run/Orthogroups.GeneCount.tsv",header=T, row.names=1))
#rownames(ogdf)<-unlist(lapply(rownames(ogdf),FUN=function(x) strsplit(x,"_")[[1]][1]))
rownames(ogdf)[rownames(ogdf)=="Limulus.genome"]<-"Limulus_genome"
rownames(ogdf)[rownames(ogdf)=="Hypochithonius.genome"]<-"Hypochithonius_genome"

#rownames(ogdf)<-Hmisc::upFirst(rownames(ogdf))
rownames(ogdf)[!(rownames(ogdf) %in% tr1$tip.label)]
ogdf<-ogdf[(rownames(ogdf) %in% tr1$tip.label),]
tr1$tip.label[!(tr1$tip.label %in% rownames(ogdf) )]

tr8$tip.label[!(tr8$tip.label %in% rownames(ogdf) )]

save(tr1,tr2, tr3,tr4,tr5,tr6,tr0,tr7,tr8,tr9,tr10,ogdf,nodenames1,nodenames2,nodenames3,nodenames4,nodenames5,nodenames6,nodenames0,nodenames7,nodenames8,nodenames9,nodenames10,treelist,file="Chelicerate_og4dollo.Rdata")
#2 run Dollo on server
}else{ load("Chelicerate_og4dollo.Rdata")}

#set aside OG which only exist in one species
ogdf_single<-ogdf %>% data.frame() %>% dplyr::select(where(~ sum(.>0)==1))
ogdf2<-ogdf %>% data.frame() %>% dplyr::select(where(~ sum(.>0)>=2))

##Unassigned Genes from Orthofinder
##Represent OG only found in single tip (==gained at tip)

#process tsv file:
#cat Orthogroups_UnassignedGenes.tsv | perl -p -e 's/\t+/,/g' | tail -n+2  > Orthogroups_UnassignedGenes.csv
undat<-read.csv("../july_run/Orthogroups_UnassignedGenes.csv",header=F)
undat$V2<-unlist(lapply(undat$V2,function(x) paste(strsplit(x,"_")[[1]][1])))
undat$V2<-unlist(lapply(undat$V2,function(x) paste(strsplit(x,"\\|")[[1]][1])))
#hold off on parsing tips, sequences IDs too ambiguous to link to orthofinder table
#tipOGgains<-table(undat$V2)
#tipgainlist<-as.list(unstack(undat,undat[[1]]~undat[[2]]))

```


```{r dollo_plots_function}
# plot_Dollo_results()
# TREE<-tr0
# DOLLO.RESULTS<-"dolloASR_results_H0.Rds"
# LOSS.GAIN.PRES.ABS<-"loss_gain_pres_abs_H0.Rds"
# HYPOTHESISNAME<-treelist["tr0"]
# NODENAMES<-nodenames0
plot_Dollo_results<-function(TREE,DOLLO.RESULTS,LOSS.GAIN.PRES.ABS,HYPOTHESISNAME,NODENAMES){
#3 visualize results for H1
load(DOLLO.RESULTS)
tree<-TREE
nodenames<-NODENAMES
pres.og<-rowSums(dollo.anc.states)
abs.og<-(ncol(dollo.anc.states)-pres.og)
dollo.mat<-cbind(abs.og,pres.og)

#infer gains and losses from presence/absence
load(LOSS.GAIN.PRES.ABS)
gainloss.dollo.mat<-t(apply(lossgain.table,MARGIN=1, function(x) table(factor(x, levels=c("a","g","l","p"))))) %>% data.frame()
rownames(gainloss.dollo.mat)<-c(tree$tip.label,nodenames)
# a=absent & not present in ancestor
# p=inherited from ancestor (previous node)
# g=gain compared to previous node
# l=lost from previous node

# gainloss.dollo.mat is ordered tips, then nodes
# each rows reflect outcomes (ie what happened btween ancestor and tip is shown on tip's row), 
# so no gains or losses can be inferred at root
gainloss.dollo.mat.nodes<-gainloss.dollo.mat[nodenames,]
#root is in table, but no gain/loss shown as would require comparison with earlier node

dm<-dollo.mat %>% as.data.frame %>% tibble::rownames_to_column()

gl<-gainloss.dollo.mat.nodes %>% as.data.frame %>% tibble::rownames_to_column()
combined<-left_join(dm,gl) %>% dplyr::select(pres.og,abs.og,g,l,a,p)
rownames(combined)<-rownames(dollo.mat)
gainloss.dollo.mat.nodes<-combined

 p<-ggtree(tree,right=T,branch.length="none") +geom_tiplab(aes(fontface=3),size=1.5)+ geom_label2(label=nodenames, aes(subset= !isTip),hjust=1.2,size = 1.5) + hexpand(0.2) +hexpand(-1)
  pies <- nodepie(gainloss.dollo.mat.nodes %>%  mutate(node=Ntip(tree)+1:Nnode(tree)),cols=1:2)
  pies3 <- lapply(pies, function(g) g+scale_fill_manual(values = c(absent.col, present.col)))

  p_abspres<-p + ggpp::geom_plot(data=td_filter(!isTip), mapping=aes(x=x,y=y, label=pies3), vp.width=0.035, vp.height=0.035, hjust=0.5, vjust=0.5) 
 #setup dummy plot to borrow legend for pie chart colors
  e1 <- ggplot(gainloss.dollo.mat.nodes %>%  dplyr::select(pres.og,abs.og) %>%
    tibble::rownames_to_column()%>% pivot_longer(-rowname), aes(x=rowname, value, fill=name)) +
    geom_bar(stat="identity", width=1) + theme(legend.background = element_rect(color = NA, fill = NA))+ 
    scale_fill_manual(name="",values=c(absent.col,present.col),labels=c("absent","present"))
 #leg<-cowplot::get_legend(e1) #works but gives annoying warning
 leg1<-cowplot::get_plot_component(e1, "guide-box", return_all = TRUE)
 p_abspres<- p_abspres + annotation_custom(grob = leg1[[1]][[1]], xmin =  0, xmax = 0.5, ymin = 1, ymax = 5) 
 
 p<-ggtree(tree,right=T,branch.length="none") +geom_tiplab(aes(fontface=3),size=1.5)+ geom_label2(label=nodenames, aes(subset= !isTip),hjust=1.2,size = 1.5) + hexpand(0.2) +hexpand(-1)
  pies <- nodepie(gainloss.dollo.mat.nodes %>%  mutate(node=Ntip(tree)+1:Nnode(tree)),cols=3:6)
  pies4 <- lapply(pies, function(g) g+scale_fill_manual(values = c(col.a,col.g,col.l,col.p)))
  p_gainloss<-p + ggpp::geom_plot(data=td_filter(!isTip), mapping=aes(x=x,y=y, label=pies4), vp.width=0.035, vp.height=0.035, hjust=0.5, vjust=0.5) 
 #setup dummy plot to borrow legend for pie chart colors
  e1 <- ggplot(gainloss.dollo.mat.nodes %>%  dplyr::select(a,g,l,p) %>%
    tibble::rownames_to_column()%>% pivot_longer(-rowname), aes(x=rowname, value, fill=name)) +
    geom_bar(stat="identity", width=1) + theme(legend.background = element_rect(color = NA, fill = NA))+ 
    scale_fill_manual(name="",values=c(col.a,col.g,col.l,col.p),labels=c("absent","gains","losses","retained"))
 leg1<-cowplot::get_plot_component(e1, "guide-box", return_all = TRUE)
 p_gainloss<- p_gainloss + annotation_custom(grob = leg1[[1]][[1]], xmin =  0, xmax = 0.5, ymin = 1, ymax = 5) 
 

#tidy up table of node states
#finalord<-c(23,43,28,29,27,30,31,37,32,33,34,35,36,26,38,25,24,42,39,40,41)
#initalord<-23:43
#matchord<-match(finalord,initalord)
nice<-gainloss.dollo.mat.nodes
colnames(nice)<-c("OGs_present","OGs_absent","Gains","Losses","Absent","Retained")
nice2<-nice %>% tibble::rownames_to_column("Node") %>% bind_rows(summarise(., across(where(is.numeric), sum), across(!where(is.numeric), ~"Totals")))

cp1<-cowplot::plot_grid(p_abspres,p_gainloss,nrow=1,ncol=2, labels=c("A","B"))
#cowplot::ggdraw(cp1)
plot_save<-cowplot::ggdraw(cowplot::add_sub(cp1,paste0("OrthoFinder OG dynamics under Dollo parsimony under hypothesis:",HYPOTHESISNAME,"\n Porportion of all OGs inferred as present at each ancestor under Dollo (A); ancestral OGs were further compared to immediate ancestral nodes to determine \nthe proportion of the  ",ncol(ogdf), " orthogroups which constitue gains, losses, ancestral retention and ancestral absence (B)."),x=.01,hjust=0,size=12))
return(list(plot_save,nice2))
}
```

```{r plot_results}
h0<-plot_Dollo_results(TREE=tr0, DOLLO.RESULTS="dolloASR_results_H0.Rds", LOSS.GAIN.PRES.ABS="loss_gain_pres_abs_H0.Rds", HYPOTHESISNAME=treelist["tr0"], NODENAMES=nodenames0)
h1<-plot_Dollo_results(TREE=tr1, DOLLO.RESULTS="dolloASR_results_H1.Rds", LOSS.GAIN.PRES.ABS="loss_gain_pres_abs_H1.Rds", HYPOTHESISNAME=treelist["tr01"], NODENAMES=nodenames1)
h2<-plot_Dollo_results(TREE=tr2, DOLLO.RESULTS="dolloASR_results_H2.Rds", LOSS.GAIN.PRES.ABS="loss_gain_pres_abs_H2.Rds", HYPOTHESISNAME=treelist["tr02"], NODENAMES=nodenames2)
h3<-plot_Dollo_results(TREE=tr3, DOLLO.RESULTS="dolloASR_results_H3.Rds", LOSS.GAIN.PRES.ABS="loss_gain_pres_abs_H3.Rds", HYPOTHESISNAME=treelist["tr03"], NODENAMES=nodenames3)
h4<-plot_Dollo_results(TREE=tr4, DOLLO.RESULTS="dolloASR_results_H4.Rds", LOSS.GAIN.PRES.ABS="loss_gain_pres_abs_H4.Rds", HYPOTHESISNAME=treelist["tr04"], NODENAMES=nodenames4)
h5<-plot_Dollo_results(TREE=tr5, DOLLO.RESULTS="dolloASR_results_H5.Rds", LOSS.GAIN.PRES.ABS="loss_gain_pres_abs_H5.Rds", HYPOTHESISNAME=treelist["tr05"], NODENAMES=nodenames5)
h6<-plot_Dollo_results(TREE=tr6, DOLLO.RESULTS="dolloASR_results_H6.Rds", LOSS.GAIN.PRES.ABS="loss_gain_pres_abs_H6.Rds", HYPOTHESISNAME=treelist["tr06"], NODENAMES=nodenames6)
h7<-plot_Dollo_results(TREE=tr7, DOLLO.RESULTS="dolloASR_results_H7.Rds", LOSS.GAIN.PRES.ABS="loss_gain_pres_abs_H7.Rds", HYPOTHESISNAME=treelist["tr07"], NODENAMES=nodenames7)
h8<-plot_Dollo_results(TREE=tr8, DOLLO.RESULTS="dolloASR_results_H8.Rds", LOSS.GAIN.PRES.ABS="loss_gain_pres_abs_H8.Rds", HYPOTHESISNAME=treelist["tr08"], NODENAMES=nodenames8)
h9<-plot_Dollo_results(TREE=tr9, DOLLO.RESULTS="dolloASR_results_H9.Rds", LOSS.GAIN.PRES.ABS="loss_gain_pres_abs_H9.Rds", HYPOTHESISNAME=treelist["tr09"], NODENAMES=nodenames9)
h10<-plot_Dollo_results(TREE=tr10, DOLLO.RESULTS="dolloASR_results_H10.Rds", LOSS.GAIN.PRES.ABS="loss_gain_pres_abs_H10.Rds", HYPOTHESISNAME=treelist["tr10"], NODENAMES=nodenames10)

```
### OrthoFinder OG dynamics under Dollo parsimony

```{r pdfplot, eval=F}
pdf("Dollo_trees_allOG_twomore.pdf",width=12, height=10, useDingbats = F)
cowplot::ggdraw(h0[[1]])
cowplot::ggdraw(h1[[1]])
cowplot::ggdraw(h2[[1]])
cowplot::ggdraw(h3[[1]])
cowplot::ggdraw(h4[[1]])
cowplot::ggdraw(h5[[1]])
cowplot::ggdraw(h6[[1]])
cowplot::ggdraw(h7[[1]])
cowplot::ggdraw(h8[[1]])
cowplot::ggdraw(h9[[1]])
cowplot::ggdraw(h10[[1]])
dev.off()
#pdf("Dollo_trees_twomore.pdf",width=12, height=10, useDingbats = F)
#cowplot::ggdraw(h9[[1]])
#cowplot::ggdraw(h10[[1]])
#dev.off()
#save tables
library(writexl)
 list_of_dfs <- list(h0[[2]],h1[[2]],h2[[2]],h3[[2]],h4[[2]],h5[[2]],h6[[2]],h7[[2]],h8[[2]],h9[[2]],h10[[2]])
 names(list_of_dfs)<-gsub("Constraint_","",treelist)[order(names(treelist))]
 write_xlsx(list_of_dfs, "Aranchnid_Dollo_results.05.22.xlsx")
#
 #list_of_dfs <- list(h9[[2]],h10[[2]])
# names(list_of_dfs)<-gsub("Constraint_","",treelist)[order(names(treelist))][10:11]
# write_xlsx(list_of_dfs, "Aranchnid_Dollo_results.05.22.xlsx")
```


#### Fisher's exact test
Are significantly more losses required under either topology?
```{r ttests,include=T}
mat<-matrix(c(tail(h0[[2]]$Retained,1),tail(h1[[2]]$Retained,1),tail(h0[[2]]$Losses,1),tail(h1[[2]]$Losses,1)),2,2)
pval<-fisher.test(mat)$p
oddsr<-fisher.test(mat)$estimate
#tail(nice1$Losses,1)/tail(nice1$Retained,1)
#tail(nice2$Losses,1)/tail(nice2$Retained,1)
#tail(nice2$Losses,1)/tail(nice1$Losses,1)
dimnames(mat)<-list(c(treelist["tr0"],treelist["tr01"]),c("retained","lost"))
knitr::kable(mat)
perc_inc<-100*(tail(h0[[2]]$Losses,1)-tail(h1[[2]]$Losses,1))/tail(h0[[2]]$Losses,1)


xtab<-as.table(rbind(
  c(tail(h0[[2]]$Retained,1),tail(h1[[2]]$Retained,1),tail(h2[[2]]$Retained,1),tail(h3[[2]]$Retained,1),tail(h4[[2]]$Retained,1),tail(h5[[2]]$Retained,1),tail(h6[[2]]$Retained,1),tail(h7[[2]]$Retained,1),tail(h8[[2]]$Retained,1),tail(h9[[2]]$Retained,1),tail(h10[[2]]$Retained,1)),
  c(tail(h0[[2]]$Losses,1),tail(h1[[2]]$Losses,1),tail(h2[[2]]$Losses,1),tail(h3[[2]]$Losses,1),tail(h4[[2]]$Losses,1),tail(h5[[2]]$Losses,1),tail(h6[[2]]$Losses,1),tail(h7[[2]]$Losses,1),tail(h8[[2]]$Losses,1),tail(h9[[2]]$Losses,1),tail(h10[[2]]$Losses,1))))
dimnames(xtab)<-list(event=c("Retained","Lost"),hypothesis=treelist[c("tr0","tr01","tr02","tr03","tr04","tr05","tr06","tr07","tr08","tr09","tr10")])
fishertests<-rstatix::pairwise_fisher_test(xtab)
knitr::kable(t(xtab))

#add fisher results to spreadsheets
library(writexl)
 list_of_dfs <- list(h0[[2]],h1[[2]],h2[[2]],h3[[2]],h4[[2]],h5[[2]],h6[[2]],h7[[2]],h8[[2]],h9[[2]],h10[[2]],fishertests)
 names(list_of_dfs)<-c(gsub("Constraint_","",treelist)[order(names(treelist))],"Fishers_pairwise")
 write_xlsx(list_of_dfs, "Aranchnid_Dollo_results.05.22.xlsx")

```
Fisher's exact test finds a significant difference in the odds ratio, p=`r pval`, with ~`r perc_inc`% more orthogroup losses required across the `r `treelist["tr0"]` tree than in the `r treelist["tr01"]` tree.


#### Repeat analysis using only Orthogroups shared between at least two species 
No single-species orthogroups
````{r tree_bars}
tree<-tr0
nodenames<-nodenames0

#plot species-unique OG as bars
g<-ggtree(tree,right=T,branch.length="none") +geom_tiplab(aes(fontface=3),size=1.5) + geom_label2(label=nodenames, aes(subset= !isTip),hjust=1.2,size = 2) + hexpand(0.5) + hexpand(0.25, -1)+ theme(legend.position = "none")
shared<-ogdf2 %>% tibble::rownames_to_column("tip") %>% pivot_longer(-tip,names_to="OG",values_to="count") %>% group_by(tip) %>% summarize(shared_across=sum(count))
singles<-ogdf_single %>% tibble::rownames_to_column("tip") %>% pivot_longer(-tip,names_to="OG",values_to="count") %>% group_by(tip) %>% summarize(single_species=sum(count))
tots<-left_join(shared,singles) %>% pivot_longer(-tip,names_to = "totaltype", values_to = "value")

p1 <- ggplot(tots,aes(x=tip,fill=totaltype,y=value))  + geom_bar(stat="identity")+ labs(x="OrthoGroups detected by species",fill="Number of Orthogroups") + 
    #geom_text(aes(label=total, y= total+100),hjust=0) +
    coord_flip() + theme_tree2() 
library(aplot)
tree_single_bars<-p1 %>% insert_left(g, width=2) 
#plotted with shared-only dollo trees below

```


```{r dollo_plot_function2}
# plot_Dollo_results()
# TREE<-tr0
# DOLLO.RESULTS<-"dolloASR_results_H0.Rds"
# LOSS.GAIN.PRES.ABS<-"loss_gain_pres_abs_H0.Rds"
# HYPOTHESISNAME<-treelist["tr0"]
# NODENAMES<-nodenames0
plot_Dollo_results2<-function(TREE,DOLLO.RESULTS,LOSS.GAIN.PRES.ABS,HYPOTHESISNAME,NODENAMES){
#3 visualize results for H1
load(DOLLO.RESULTS)
tree<-TREE
nodenames<-NODENAMES

dollo.anc.states_shared<- dollo.anc.states %>% dplyr::select(colnames(ogdf2))
pres.og<-rowSums(dollo.anc.states_shared)
abs.og<-(ncol(dollo.anc.states_shared)-pres.og)
dollo.mat<-cbind(abs.og,pres.og)

#infer gains and losses from presence/absence
load(LOSS.GAIN.PRES.ABS)
lossgain.table_single<-lossgain.table %>%  dplyr::select(colnames(ogdf_single))
lossgain.table_shared<-lossgain.table %>%  dplyr::select(colnames(ogdf2))

gainloss.dollo.mat<-t(apply(lossgain.table_shared,MARGIN=1, function(x) table(factor(x, levels=c("a","g","l","p"))))) %>% data.frame()
rownames(gainloss.dollo.mat)<-c(tree$tip.label,nodenames)

# a=absent & not present in ancestor
# p=inherited from ancestor (previous node)
# g=gain compared to previous node
# l=lost from previous node

# gainloss.dollo.mat is ordered tips, then nodes
# each rows reflect outcomes (ie what happened btween ancestor and tip is shown on tip's row), 
# so no gains or losses can be inferred at root
gainloss.dollo.mat.nodes<-gainloss.dollo.mat[nodenames,]
#root is in table, but no gain/loss shown as would require comparison with earlier node

dm<-dollo.mat %>% as.data.frame %>% tibble::rownames_to_column()
gl<-gainloss.dollo.mat.nodes %>% as.data.frame %>% tibble::rownames_to_column()
combined<-left_join(dm,gl) %>% dplyr::select(pres.og,abs.og,g,l,a,p)
rownames(combined)<-rownames(dollo.mat)
gainloss.dollo.mat.nodes<-combined


 p<-ggtree(tree,right=T,branch.length="none") +geom_tiplab(aes(fontface=3),size=1.5) + geom_label2(label=nodenames, aes(subset= !isTip),hjust=1.2,size = 1.5) + hexpand(0.2)+hexpand(-1)
  pies <- nodepie(gainloss.dollo.mat.nodes %>%  mutate(node=Ntip(tr1)+1:Nnode(tr1)),cols=1:2)
  pies1 <- lapply(pies, function(g) g+scale_fill_manual(values = c(absent.col, present.col)))
  p_abspres<-p + ggpp::geom_plot(data=td_filter(!isTip), mapping=aes(x=x,y=y, label=pies1), vp.width=0.035, vp.height=0.035, hjust=0.5, vjust=0.5) 
 #setup dummy plot to borrow legend for pie chart colors
  e1 <- ggplot(gainloss.dollo.mat.nodes %>%  dplyr::select(pres.og,abs.og) %>%
    tibble::rownames_to_column()%>% pivot_longer(-rowname), aes(x=rowname, value, fill=name)) +
    geom_bar(stat="identity", width=1) + theme(legend.background = element_rect(color = NA, fill = NA))+
    scale_fill_manual(name="",values=c(absent.col,present.col),labels=c("absent","present"))
 leg1<-cowplot::get_plot_component(e1, "guide-box", return_all = TRUE)
 p_abspres<-p_abspres + annotation_custom(grob = leg1[[1]][[1]], xmin = -0, xmax = 0.5, ymin = 10, ymax = 15) 

 
 p<-ggtree(tree,right=T,branch.length="none") +geom_tiplab(aes(fontface=3),size=1.5) + geom_label2(label=nodenames, aes(subset= !isTip),hjust=1.2,size = 1.5)+ hexpand(0.2) +hexpand(-1)
  pies <- nodepie(gainloss.dollo.mat.nodes %>%  mutate(node=Ntip(tr1)+1:Nnode(tr1)),cols=3:6)
  pies2 <- lapply(pies, function(g) g+scale_fill_manual(values = c(col.a,col.g,col.l,col.p)))
  p_gainloss<-p + ggpp::geom_plot(data=td_filter(!isTip), mapping=aes(x=x,y=y, label=pies2), vp.width=0.035, vp.height=0.035, hjust=0.5, vjust=0.5) 
 #setup dummy plot to borrow legend for pie chart colors
  e1 <- ggplot(gainloss.dollo.mat.nodes %>%  dplyr::select(a,g,l,p) %>%
    tibble::rownames_to_column()%>% pivot_longer(-rowname), aes(x=rowname, value, fill=name)) +
    geom_bar(stat="identity", width=1) + theme(legend.background = element_rect(color = NA, fill = NA))+ 
    scale_fill_manual(name="",values=c(col.a,col.g,col.l,col.p),labels=c("absent","gains","losses","retained"))
 leg1<-cowplot::get_plot_component(e1, "guide-box", return_all = TRUE)
 p_gainloss<- p_gainloss + annotation_custom(grob = leg1[[1]][[1]], xmin =  0, xmax = 0.5, ymin = 10, ymax = 15) 
 

#tidy up table of node states
#finalord<-c(23,43,28,29,27,30,31,37,32,33,34,35,36,26,38,25,24,42,39,40,41)
#initalord<-23:43
#matchord<-match(finalord,initalord)
#nice<-gainloss.dollo.mat.nodes[matchord,]
nice<-gainloss.dollo.mat.nodes
colnames(nice)<-c("OGs_present","OGs_absent","Gains","Losses","Absent","Retained")
nice1<-nice %>% tibble::rownames_to_column("Node") %>% bind_rows(summarise(., across(where(is.numeric), sum), across(!where(is.numeric), ~"Totals")))

cp1<-cowplot::plot_grid(p_abspres,p_gainloss,nrow=1,ncol=2, labels=c("A","B"))
#cowplot::ggdraw(cp1)
plot_save<-cowplot::ggdraw(cowplot::add_sub(cp1,paste0("OrthoFinder OG dynamics under Dollo parsimony under hypothesis: ",HYPOTHESISNAME,"\n Proportion of all ",ncol(ogdf2), " OGs (shared between at least 2 species) inferred as present at each ancestor under Dollo are shown (A);  ancestral OGs were \nfurther compared to immediate ancestral nodes to determine the proportion of the  ",ncol(ogdf2), " shared orthogroups which constitue gains, losses, \nancestral retention and ancestral absence (B)."),x=.01,hjust=0,size=12))
return(list(plot_save,nice1))
}

```


```{r plot_results_shared}
h0<-plot_Dollo_results2(TREE=tr0, DOLLO.RESULTS="dolloASR_results_H0.Rds", LOSS.GAIN.PRES.ABS="loss_gain_pres_abs_H0.Rds", HYPOTHESISNAME=treelist["tr0"], NODENAMES=nodenames0)
h1<-plot_Dollo_results2(TREE=tr1, DOLLO.RESULTS="dolloASR_results_H1.Rds", LOSS.GAIN.PRES.ABS="loss_gain_pres_abs_H1.Rds", HYPOTHESISNAME=treelist["tr01"], NODENAMES=nodenames1)
h2<-plot_Dollo_results2(TREE=tr2, DOLLO.RESULTS="dolloASR_results_H2.Rds", LOSS.GAIN.PRES.ABS="loss_gain_pres_abs_H2.Rds", HYPOTHESISNAME=treelist["tr02"], NODENAMES=nodenames2)
h3<-plot_Dollo_results2(TREE=tr3, DOLLO.RESULTS="dolloASR_results_H3.Rds", LOSS.GAIN.PRES.ABS="loss_gain_pres_abs_H3.Rds", HYPOTHESISNAME=treelist["tr03"], NODENAMES=nodenames3)
h4<-plot_Dollo_results2(TREE=tr4, DOLLO.RESULTS="dolloASR_results_H4.Rds", LOSS.GAIN.PRES.ABS="loss_gain_pres_abs_H4.Rds", HYPOTHESISNAME=treelist["tr04"], NODENAMES=nodenames4)
h5<-plot_Dollo_results2(TREE=tr5, DOLLO.RESULTS="dolloASR_results_H5.Rds", LOSS.GAIN.PRES.ABS="loss_gain_pres_abs_H5.Rds", HYPOTHESISNAME=treelist["tr05"], NODENAMES=nodenames5)
h6<-plot_Dollo_results2(TREE=tr6, DOLLO.RESULTS="dolloASR_results_H6.Rds", LOSS.GAIN.PRES.ABS="loss_gain_pres_abs_H6.Rds", HYPOTHESISNAME=treelist["tr06"], NODENAMES=nodenames6)
h7<-plot_Dollo_results2(TREE=tr7, DOLLO.RESULTS="dolloASR_results_H7.Rds", LOSS.GAIN.PRES.ABS="loss_gain_pres_abs_H7.Rds", HYPOTHESISNAME=treelist["tr07"], NODENAMES=nodenames7)
h8<-plot_Dollo_results2(TREE=tr8, DOLLO.RESULTS="dolloASR_results_H8.Rds", LOSS.GAIN.PRES.ABS="loss_gain_pres_abs_H8.Rds", HYPOTHESISNAME=treelist["tr08"], NODENAMES=nodenames8)
h9<-plot_Dollo_results2(TREE=tr9, DOLLO.RESULTS="dolloASR_results_H9.Rds", LOSS.GAIN.PRES.ABS="loss_gain_pres_abs_H9.Rds", HYPOTHESISNAME=treelist["tr09"], NODENAMES=nodenames9)
h10<-plot_Dollo_results2(TREE=tr10, DOLLO.RESULTS="dolloASR_results_H10.Rds", LOSS.GAIN.PRES.ABS="loss_gain_pres_abs_H10.Rds", HYPOTHESISNAME=treelist["tr10"], NODENAMES=nodenames10)

```
### OrthoFinder OG dynamics under Dollo parsimony

```{r pdfplot_shared, eval=F}
pdf("Dollo_trees_sharedOG.pdf",width=12, height=10, useDingbats = F)

cowplot::ggdraw(cowplot::add_sub(ggplotify::as.grob(tree_single_bars),label="Arachnida constraint tree with total orthogroups mapped.  Orthogroups shared by more than one species (red) are distinguished from orthogroups \ndetected in a single taxon (blue). Note: Unassigned OGs (single sequences) have not been included.",x=0.1,hjust=0,size=8))

cowplot::ggdraw(h0[[1]])
cowplot::ggdraw(h1[[1]])
cowplot::ggdraw(h2[[1]])
cowplot::ggdraw(h3[[1]])
cowplot::ggdraw(h4[[1]])
cowplot::ggdraw(h5[[1]])
cowplot::ggdraw(h6[[1]])
cowplot::ggdraw(h7[[1]])
cowplot::ggdraw(h8[[1]])
cowplot::ggdraw(h9[[1]])
cowplot::ggdraw(h10[[1]])
dev.off()

#save tables
library(writexl)
 list_of_dfs <- list(h0[[2]],h1[[2]],h2[[2]],h3[[2]],h4[[2]],h5[[2]],h6[[2]],h7[[2]],h8[[2]],h9[[2]],h10[[2]])
 names(list_of_dfs)<-gsub("Constraint_","",treelist)[order(names(treelist))]
 write_xlsx(list_of_dfs, "Aranchnid_Dollo_results_sharedOG.5.22.xlsx")

```

```{r troubleshoot}
DOLLO.RESULTS<-"dolloASR_results_H0.Rds"
LOSS.GAIN.PRES.ABS<-"loss_gain_pres_abs_H0.Rds"
HYPOTHESISNAME<-treelist["tr0"]
NODENAMES<-nodenames0
TREE<-tr0
load(DOLLO.RESULTS)
tree<-TREE
nodenames<-NODENAMES
pres.og<-rowSums(dollo.anc.states)
abs.og<-(ncol(dollo.anc.states)-pres.og)
dollo.mat.all<-cbind(abs.og,pres.og)

dollo.anc.states_shared<- dollo.anc.states %>% dplyr::select(colnames(ogdf2))
pres.og<-rowSums(dollo.anc.states_shared)
abs.og<-(ncol(dollo.anc.states_shared)-pres.og)
dollo.mat.shared<-cbind(abs.og,pres.og)

#infer gains and losses from presence/absence
load(LOSS.GAIN.PRES.ABS)
lossgain.table_single<-lossgain.table %>%  dplyr::select(colnames(ogdf_single))
lossgain.table_shared<-lossgain.table %>%  dplyr::select(colnames(ogdf2))

```


#### Fisher's exact test
Are significantly more losses required under either topology?
```{r ttests,include=T}
mat<-matrix(c(tail(nice1$Retained,1),tail(nice2$Retained,1),tail(nice1$Losses,1),tail(nice2$Losses,1)),2,2)
pval<-fisher.test(mat)$p
oddsr<-fisher.test(mat)$estimate
#tail(nice1$Losses,1)/tail(nice1$Retained,1)
#tail(nice2$Losses,1)/tail(nice2$Retained,1)
#tail(nice2$Losses,1)/tail(nice1$Losses,1)
dimnames(mat)<-list(c("constraint","nonconstraint"),c("retained","lost"))
knitr::kable(mat)
perc_inc<-100*(tail(nice1$Losses,1)-tail(nice2$Losses,1))/tail(nice1$Losses,1)


xtab<-as.table(rbind(c(tail(h0[[2]]$Retained,1),tail(h1[[2]]$Retained,1),tail(h2[[2]]$Retained,1),tail(h3[[2]]$Retained,1),tail(h4[[2]]$Retained,1),tail(h5[[2]]$Retained,1),tail(h6[[2]]$Retained,1),tail(h7[[2]]$Retained,1),tail(h8[[2]]$Retained,1),tail(h9[[2]]$Retained,1),tail(h10[[2]]$Retained,1)),c(tail(h0[[2]]$Losses,1),tail(h1[[2]]$Losses,1),tail(h2[[2]]$Losses,1),tail(h3[[2]]$Losses,1),tail(h4[[2]]$Losses,1),tail(h5[[2]]$Losses,1),tail(h6[[2]]$Losses,1),tail(h7[[2]]$Losses,1),tail(h8[[2]]$Losses,1),tail(h9[[2]]$Losses,1),tail(h10[[2]]$Losses,1))))
dimnames(xtab)<-list(event=c("Retained","Lost"),hypothesis=treelist[c("tr0","tr01","tr02","tr03","tr04","tr05","tr06","tr07","tr08","tr09","tr10")])
fishertests<-rstatix::pairwise_fisher_test(xtab)
knitr::kable(t(xtab))
group1_lost_to_retained<-group2_lost_to_retained<-xtab["Lost",]/xtab["Retained",]

str(group1_lost_to_retained)
fishertests<-fishertests %>% left_join(., data.frame(group1_lost_to_retained, group1 = names(group1_lost_to_retained)))  %>% left_join(., data.frame(group2_lost_to_retained, group2 = names(group2_lost_to_retained))) %>% mutate(diff=group1_lost_to_retained-group2_lost_to_retained) 
#to get symmetric matrix
fishertests[c("group1", "group2")] <- lapply(fishertests[c("group1", "group2")], factor, levels=unique(unlist(fishertests[c("group1", "group2")]))) 

xtabs(diff ~ group1 + group2, data=fishertests)
fishertests2<-fishertests %>% mutate(diff_signif=paste0(signif(diff,4),p.adj.signif)) %>% mutate(diff_signif2=if_else(p.adj.signif=="ns",NA,signif(diff,4)))

dm<-xtabs(diff ~ group1 + group2, data=fishertests)
dm + t(dm) 
dm2<-xtabs(diff_signif2 ~ group1 + group2, data=fishertests2,na.action = na.pass)
signif_dm<-dm2 + t(dm2)
signif_dm<-as.data.frame.matrix(signif_dm)
signif_dm<-cbind(names(signif_dm),signif_dm);names(signif_dm)[1]<-"topology"
#add fisher results to spreadsheets
library(writexl)
 list_of_dfs <- list(h0[[2]],h1[[2]],h2[[2]],h3[[2]],h4[[2]],h5[[2]],h6[[2]],h7[[2]],h8[[2]],h9[[2]],h10[[2]],fishertests,signif_dm)
 names(list_of_dfs)<-c(gsub("Constraint_","",treelist)[order(names(treelist))],"Fishers_results","Fishers_pairwise")
 write_xlsx(list_of_dfs, "Aranchnid_Dollo_results_sharedOG.05.27.xlsx")

```
Fisher's exact test finds a significant difference in the odds ratio, p=`r pval`, with ~`r perc_inc`% more orthogroup losses required across the constraint tree than in the non-constraint tree.


#### Orthogroups shared among tips
```{r upgma}

dm <- dist(ogdf2)
upgmatree <- upgma(dm)  %>% reroot(., node.number=getMRCA(.,tip=c("Epiperipatus_transcriptome","Opisthopatus_transcriptome")),position=0.1)%>% ladderize()
plot(upgmatree)
plot(tr1)
# limit to chelicerata 
keep<-tr1$tip.label[unlist(Descendants(tr1,node=getMRCA(.,tip=c("Argiope_genome","Pycnogonum_genome")),type="tips"))]
ogdf_chel<-ogdf2[rownames(ogdf2) %in% keep,]
dm <- dist(ogdf_chel)
upgmatree <- upgma(dm)  %>% reroot(., node.number=getMRCA(.,tip=c("Pycnogonum_genome","Anoplo_transcriptome")),position=0.1)%>% ladderize()
plot(upgmatree)

#plot UPGMa tree and heatmap of distances
  dist_df <- as.data.frame(as.matrix(dm))
 longdf<-dist_df %>% tibble::rownames_to_column("label")  %>% pivot_longer(!label,names_to="comparison",values_to="distance")
longdf$label<-factor(longdf$label,levels=upgmatree$tip.label)
longdf$comparison<-factor(longdf$comparison,levels=upgmatree$tip.label)

ggplot(longdf, aes(x = label, y = comparison, fill = distance)) +
  geom_tile() +
  scale_fill_gradient(low = "white", high = "red") + # Customize colors as needed
  labs(title = "Heatmap of Distances from hclust", x = "Sample", y = "Sample") +
  theme_minimal() + # Minimal theme for a clean look
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) # Rotate x-axis labels if necessary

ogdfPA<-ogdf2 %>%  mutate(across(where(is.numeric), ~ ifelse(. > 0, 1, .)))
write.table(ogdfPA,file="Orthogroups_shared.Incidence.tsv")
#make heatmap of co-occurences, use order from UPGMA 
dm <- dist(ogdfPA)
upgmatree <- upgma(dm)  %>% reroot(., node.number=getMRCA(.,tip=c("Epiperipatus_transcriptome","Opisthopatus_transcriptome")),position=0.1)%>% ladderize()
plot(upgmatree)
hclust<-hclust(dm)


dat<- ogdf2 %>%  tibble::rownames_to_column("label")  %>% pivot_longer(!label,names_to="OG",values_to="present") %>% filter(present>0) %>%data.frame()
#out <- crossprod(table(long_ogdf2[1:2]))
library("Matrix")

# factors for indexing matrix entries and naming dimensions
trx.fac <- factor(dat[,2])
itm.fac <- factor(dat[,1])

s <- sparseMatrix(
        as.numeric(trx.fac), 
        as.numeric(itm.fac),
        dimnames = list(
                as.character(levels(trx.fac)), 
                as.character(levels(itm.fac))),
        x = 1)

# calculating co-occurrences
v <- t(s) %*% s
dat2 <-
  v %>% as.matrix() %>% as.data.frame() %>%
  as_tibble() %>% mutate(rows=colnames(.)) %>%
  pivot_longer(-rows, names_to = "columns", values_to = "value") 
dat2$rows<-factor(dat2$rows,levels=upgmatree$tip.label)
dat2$columns<-factor(dat2$columns,levels=upgmatree$tip.label)


ggplot(dat2, aes(rows, columns)) +
  geom_tile(aes(fill = value)) +
  geom_text(aes(label = round(value, 1)),size=1) +
  scale_fill_gradient(low = "white", high = "red") +
   theme(axis.text.x = element_text(angle = 45, hjust = 1,size=4))
cooccurmat<-v %>% as.matrix() %>% as.data.frame() 
rownames(cooccurmat)<-colnames(cooccurmat)
cooccurmat<-cooccurmat[hclust$labels[rev(hclust$order)],hclust$labels[rev(hclust$order)]]
write.csv(cooccurmat,file="cooccuringOGmatrix.5.27.csv")

#devtools::install_github("nicolash2/ggdendroplot")
library(ggdendroplot)
dm <- hmReady( as.matrix(cooccurmat), colclus=hclust, rowclus=hclust ) #we use our column clustering for both row and col clustering

hclust_plot<-ggplot() + 
  geom_tile(data=dm, aes(x=x, y=y, fill=value)) +                      #heatmap
  scale_fill_gradient(low = "white", high = "red") +          #options for heatmap
  geom_dendro(hclust, ylim=c(95, 100)) +                             #upper dendrogram
  geom_dendro(hclust, xlim=c(95,100), pointing="side") +             #side dendrogram
  theme_hm()+                                                          #design
  theme(axis.title=element_blank(),axis.text.x = element_text(angle = 45, hjust = 1,size=4),axis.text.y = element_text(size=4))   


pdf("OG_co_occurence_clustered.pdf",width=12, height=10, useDingbats = F)
cowplot::ggdraw(cowplot::add_sub(hclust_plot,paste0("Hierarchical clustering of species based on presence/absence of ",ncol(ogdf2)," shared OGs."),x=0.1,hjust=0,size=8))
dev.off()



#consolidate to 12 clades
cladedf<-read.csv("../july_run/clades.csv",header=T)
ogdf_clade<-ogdf2 %>% tibble::rownames_to_column("tip") %>%left_join(.,cladedf[,1:2]) %>% group_by(Clade) %>% summarize(across(where(is.numeric),sum))
ogdfPA_clade<-ogdf_clade %>%  mutate(across(where(is.numeric), ~ ifelse(. > 0, 1, .)))
write.table(ogdfPA_clade,file="Orthogroups_clade.Incidence.tsv")
#make heatmap of co-occurences, use order from UPGMA 
dm <- ogdfPA_clade %>% data.frame(.,row.names=ogdfPA_clade$Clade) %>% dplyr::select(!Clade) %>% dist()
hclust<-hclust(dm,method="average")


dat<- ogdf_clade  %>% pivot_longer(!Clade,names_to="OG",values_to="present") %>% filter(present>0) %>%data.frame()
#out <- crossprod(table(long_ogdf2[1:2]))
library("Matrix")

# factors for indexing matrix entries and naming dimensions
trx.fac <- factor(dat[,2])
itm.fac <- factor(dat[,1])

s <- sparseMatrix(
        as.numeric(trx.fac), 
        as.numeric(itm.fac),
        dimnames = list(
                as.character(levels(trx.fac)), 
                as.character(levels(itm.fac))),
        x = 1)

# calculating co-occurrences
v <- t(s) %*% s
dat2 <-
  v %>% as.matrix() %>% as.data.frame() %>%
  as_tibble() %>% mutate(rows=colnames(.)) %>%
  pivot_longer(-rows, names_to = "columns", values_to = "value") 

cooccurmat<-v %>% as.matrix() %>% as.data.frame() 
rownames(cooccurmat)<-colnames(cooccurmat)
cooccurmat<-cooccurmat[hclust$labels[rev(hclust$order)],hclust$labels[rev(hclust$order)]]
write.csv(cooccurmat,file="cooccuringOGmatrix_clades.csv")

#devtools::install_github("nicolash2/ggdendroplot")
library(ggdendroplot)
dm <- hmReady( as.matrix(cooccurmat), colclus=hclust, rowclus=hclust ) #we use our column clustering for both row and col clustering

hclust_plot<-ggplot() + 
  geom_tile(data=dm, aes(x=x, y=y, fill=value)) +                      #heatmap
  geom_text(data=dm,aes(x=x, y=y,label = value),size=3) +
  scale_fill_gradient(low = "white", high = "red") +          #options for heatmap
  geom_dendro(hclust, ylim=c(15, 17)) +                             #upper dendrogram
 geom_dendro(hclust, xlim=c(15,17), pointing="side") +             #side dendrogram
  theme_hm()+                                                          #design
  theme(axis.title=element_blank(),axis.text.x = element_text(angle = 45, hjust = 1,size=8),axis.text.y = element_text(size=8))   


pdf("OG_co_occurence_clades.pdf",width=12, height=10, useDingbats = F)
cowplot::ggdraw(cowplot::add_sub(hclust_plot,paste0("UPGMA clustering of clades based on presence/absence of ",ncol(ogdf2)," shared OGs. Values represent number of OGs present in both clades."),x=0.1,hjust=0,size=8))
dev.off()
```
```{r,eval=FALSE}
#plotting tree and OG table (example)
 p<-ggtree(tr1,right=T) #+geom_tiplab()+ hexpand(.01)
 
  # p2
  longdf<-ogdf_chel  %>% tibble::rownames_to_column("label")  %>% 
    filter(label %in% tr1$tip.label) %>% pivot_longer(!label,names_to="OG",values_to="value")
  longdf2<-left_join(longdf,joined2 %>% filter(homo_symbol %in% sym_list[[x]]) %>% distinct(.,OG,.keep_all=T))
  
  my_breaks<-c(1,round(max(longdf2$value)/100),plyr::round_any(max(longdf2$value)/50,10),plyr::round_any(max(longdf2$value)/10,10),plyr::round_any(max(longdf2$value)/3,10),plyr::round_any(max(longdf2$value),10))
  longdf2$value[longdf2$value==0]<-NA #can specify NA color
  #set gene axis order by OG ( corresponds with overall abundance)
  p2 <- ggplot(longdf2, aes(x=reorder(homo_symbol,desc(OG)*-1), y=label)) + 
    geom_tile(aes(fill=value)) + 
    scale_fill_continuous(name = "count", trans = "log10", breaks = my_breaks, na.value = col.a, labels = c(my_breaks[1:6]),type="viridis") + 
    
     theme_minimal() + xlab(NULL) + ylab(NULL) +
    theme(axis.text=element_text(size=5),
    axis.text.x = element_text(angle = 90, hjust = 0, vjust = 0.5)) +
    scale_x_discrete(position = "top")
  p2 %>% insert_left(p1,width=0.5)
 

```
